---
title: "FanGraphs"
output: html_document
date: ''
---

<link rel="preconnect" href="https://fonts.googleapis.com">
<link rel="preconnect" href="https://fonts.gstatic.com" crossorigin>
<link href="https://fonts.googleapis.com/css2?family=Open+Sans:wght@300&family=Source+Sans+Pro&display=swap" rel="stylesheet">
<style>

body{
font-family: 'Source Sans Pro', sans-serif;
}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(readr)
library(stringr)
library(knitr)
library(kableExtra)
```

**type of data available:** game-by-game, summary statistics (split by Batting, Pitching, and Fielding)

throughout we will be using several r packages... what they are... what they do... (should this be here or in the intro section whenever we make that?)

There are two ways to use data from FanGraphs in R:

1.  Downloading the data directly from <https://www.fangraphs.com/> and importing it into R
1.  Using the [**baseballr**](https://billpetti.github.io/baseballr/index.html) package

## FanGraphs Website

The 'Leaders' and 'Teams' pages will both bring up a dashboard that can easily be downloaded. The data is very customizable. There are several tabs that will show different sets of data. And at the bottom of the page there is an option to select which variables to include in order to make a fully custom report. In the top right corner of the table it says "Export Data". The data will download as a CSV file. 

The file I chose to download is the 2022 Team Batting Stats. If the working directory is set as the source file location, you can use the `read_csv()` function (from the **readr** package) with just the file name in quotes. To check the working directory run the function `getwd`. To change the working directory go to the "Session" menu and select "Set Working Directory".

```{r, message=FALSE, warning=FALSE}
fg_download <- read_csv("FanGraphs Leaderboard.csv")
fg_download <- as.data.frame(fg_download)
```

```{r, echo=FALSE}
# row.names(fg_download) <- fg_download$Team
fg_download %>% 
  mutate(Team. = Team) %>% 
  kable() %>% 
  kable_styling("striped") %>% 
  scroll_box(width = "100%", height = "300px")
```

There are a few important things to note about the data. The variables representing walk rate and strikeout rate have '%' in the column name and in the data values. This is not ideal. Those column names must be placed in backticks ` ` whenever referenced because of the special character. Another option is to change the name using a function like `rename()`. To fix the data values use the function `str_remove()` inside of `mutate()`. We can simultaneously convert the variables to numeric instead of character with `as.numeric()`.
```{r}
# renaming the columns
fg_update <- fg_download %>% rename(BB_rate = `BB%`, K_rate = `K%`)

# fixing the data values
fg_update <- fg_update %>% 
  mutate(BB_rate = as.numeric(str_remove(BB_rate, "%")), 
         K_rate = as.numeric(str_remove(K_rate, "%")))
```


```{r, fig.align='center', echo=FALSE}
fg_update %>% 
  ggplot(aes(x = BB_rate, y = K_rate, label = Team, fill = WAR)) + 
  geom_label(alpha = 0.6, size = 3) +
  labs(title = "K% and BB% For Each MLB Team", 
       x = "Walk Rate(%)", y = "Strikeout Rate(%)") +
  scale_fill_viridis_c() +
  theme_bw()
```

Following the same process as before, I got the 2022 Pitching Leaders. Without changing any settings on FanGraphs, the data frame downloaded has 22 variables and 45 observations.  

```{r, message=FALSE, warning=FALSE}
fg_pitch_leaders <- read_csv("fg_pitch_leaders.csv")
```

```{r, echo=FALSE}
# row.names(fg_download) <- fg_download$Team
fg_pitch_leaders %>% 
  kable() %>% 
  kable_styling("striped") %>% 
  scroll_box(width = "100%", height = "300px")
```

**NOTE:** there is something weird going on with Jordan Montgomery

Again, there are variable names with strange characters in them and this is a fairly small dataset. Next we will try accessing leaderboard data, but this time using the **baseballr** package.

## baseballr

There are [functions](https://billpetti.github.io/baseballr/reference/fangraphs.html) to access game logs, leaderboards, and park factors. The game logs and leaderboards each have separate functions for pitchers and for hitters. 

```{r, message=FALSE, warning=FALSE}
library(baseballr)
```


#### Game Logs
To access game logs for pitchers: `fg_pitcher_game_logs()`
To access game logs for batters: `fg_batter_game_logs()`
To access Minor League data: `fg_milb_pitcher_game_logs()` , `fg_milb_batter_game_logs()`

Both of these functions require input of 'playerid' and 'year'. To find a player's playerid search them on the [FanGraphs](https://www.fangraphs.com/) website. The playerid is the numbers after the player's name in the URL. For example, to find Aaron Judge's playerid look at his page: <https://www.fangraphs.com/players/aaron-judge/15640/stats?position=OF>. Aaron Judge's FanGraphs player id is 15640.

```{r}
judge <- fg_batter_game_logs(playerid = 15640, year = 2021)
```

This produces a data frame with 245 variables. Each row is a different game played within the 2021 season.

```{r, message=FALSE, warning=FALSE}
judge_filtered <- judge %>% 
  dplyr::select(PlayerName, Date, Opp, Pos, AB:H, HR:BB, SO, `BB%`:BABIP, wOBA, EV:`HardHit%`)
```

```{r, message=FALSE, warning=FALSE}
judge_filtered %>% 
  sample_n(15) %>% 
  kable() %>% 
  kable_styling("striped") %>% 
  scroll_box(height = "300px")
```

```{r, fig.align='center', fig.width=7, message=FALSE, warning=FALSE, echo=FALSE}
judge_filtered %>% 
  ggplot(aes(x = EV, LA, color = factor(RBI))) +
  geom_point(alpha = 0.6) +
  labs(title = "Aaron Judge Batting", 
       x = "Exit Velocity", y = "Launch Angle", color = "RBI") +
  theme_bw()
```


#### Leaderboards
To access leaderboards for pitchers: `fg_pitcher_leaders()`
To access leaderboards for pitchers: `fg_batter_leaders()`

The arguments are 'x' for the first season of interest, 'y' for the last season of interest, 'league', and 'pitcher_type' (default is "pit" for all pitchers, other options are "sta" for starters and "rel" for relievers). There are additional optional arguments: 'ind', which specifies if if the data should look at one season or over many, and 'league', which determines if players must meet a certain qualification to be included (plate appearances for batters and innings per game for pitchers).

```{r, message=FALSE, warning=FALSE}
pitch_leaders <- fg_pitcher_leaders(x = 2020, y = 2022)
```

The resulting data frame has 303 variables. 

```{r, fig.align='center'}
pl_filtered <- pitch_leaders %>% 
  dplyr::select(playerid:G, IP:BB, SO, Pitches, AVG:BABIP, WAR:Dollars)
```

```{r, fig.align='center', fig.width=7, message=FALSE, warning=FALSE, echo=FALSE}
pl_filtered %>% 
  ggplot(aes(x = factor(Season), y = ERA)) +
  geom_jitter(alpha = 0.6, aes(color = Age),
              position=position_jitter(0.2)) +
  labs(title = "Pitchers", x = "Season") +
  scale_color_viridis_b() +
  theme_bw()
```


Accessing data from baseballr gives over double the amount of observations and hundreds of more variables.  

There are also FanGraphs functions that can scrape [park factors](https://billpetti.github.io/baseballr/reference/fg_park.html).


PROS AND CONS - I'm not sure that there's a way to cleanly format it to make a pros and cons table. Should we still put some or just not do this part?


