---
title: "Lahman"
output: html_document
date: ''
---

<link rel="preconnect" href="https://fonts.googleapis.com">
<link rel="preconnect" href="https://fonts.gstatic.com" crossorigin>
<link href="https://fonts.googleapis.com/css2?family=Open+Sans:wght@300&family=Source+Sans+Pro&display=swap" rel="stylesheet">
<style>

body{
font-family: 'Source Sans Pro', sans-serif;
}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(patchwork)
set.seed(2023)
```

The **Lahman** package contains several tables with data on pitching, hitting, fielding, and more. This package is updated with each season. The following references the 10.0-1 version. 

**type of data available:** season-by-season, broken down by team for `Batting` / `Pitching` (so if a player played on multiple teams in one season, each team as their own row), broken down by team _and_ position for `Fielding`

The main tables are:

- `People` : player names, dates of birth, death, and other biographical info
- `Batting` : batting statistics 
  - `battingStats` : calculates batting average (BA) plate appearances (PA), total bases (TB), slugging percentage (SlugPct), on-base percentage (OBP), on-base percentage + slugging (OPS), and batting average on balls in play (BABIP)
- `Pitching` : pitching statistics 
- `Fielding` : fielding statistics 
- `Salaries` : player name, year, team, league, and salary (from 1985-2016)
- `Teams` : overall season statistics
- Additional tables cover information such as All Star appearances, post season data, managers, Hall of Fame voting, awards, parks, and colleges.

A full list of the tables and their contents can be found in the [documentation](https://cran.r-project.org/web/packages/Lahman/Lahman.pdf).


```{r, message=FALSE, warning=FALSE}
library(Lahman)
```


#### People
Here is an example of the `People` table. Each row contains data about a player's birth date, hometown, death date, physical characteristics, first game, last game, and their ID for a few popular databases. 
```{r, eval=FALSE}
People %>% 
  sample_n(6)
```


```{r, echo=FALSE}
People %>% 
  sample_n(6) %>% 
  kable() %>% 
  kable_styling("striped") %>% 
  #column_spec(column = 6, width = "125px") %>% 
  scroll_box(width = "100%")
```


#### Batting 
Example: **José Iglesias**

Here is what the `Batting` table looks like for a single player. Seasons where Iglesias played for multiple teams are highlighted to help show how this table is organized.
```{r, include=FALSE}
People %>% filter(nameFirst == "Jose", nameLast == "Iglesias")
```


```{r}
iglesias <- Batting %>% 
  filter(playerID == "iglesjo01") %>% 
  arrange(desc(yearID))
```

```{r, echo=FALSE}
kable(iglesias) %>% 
  kable_styling("striped") %>% 
  row_spec(1:2, background = "skyblue") %>% 
  row_spec(9:10, background = "lightseagreen") %>% 
  scroll_box(height = "300px")
```


`battingStats()` produces a data frame with the same variables as `Batting`, and additionally calculates some statistics (as mentioned above). This table is only showing playerID, yearID, teamID, and the variables _exclusive_ to `battingStats()`.
```{r}
iglesias_stats <- battingStats(iglesias) %>% 
  dplyr::select(playerID, yearID, teamID, BA:BABIP)
```

```{r, echo=FALSE}
kable(iglesias_stats) %>% 
  kable_styling("striped") %>% 
  row_spec(1:2, background = "skyblue") %>% 
  row_spec(9:10, background = "lightseagreen") %>% 
  scroll_box(height = "300px")
```


#### Pitching 
Example: **Justin Verlander**

Just like the `Batting` table, the `Pitching` table is split by season, and team if the player pitched for several teams in one year. There are two rows for 2017 because Verlander moved from the Tigers to the Astros. 
```{r}
verlander <- Pitching %>% 
  filter(playerID == "verlaju01") %>% 
  arrange(desc(yearID))
```

```{r, echo=FALSE}
kable(head(verlander)) %>% 
  kable_styling("striped") %>% 
  row_spec(4:5, background = "lightblue") %>% 
  scroll_box(width = "100%")

# Batting %>% 
#   filter(playerID == "scherma01")
```


#### Fielding
Example: **DJ LeMahieu**

The `Fielding` table contains some of the basic statistics used for fielders. Notice the last five variables; they all say "NA". This means that when there is a missing value the Lahman database fills in "NA" as opposed to leaving the box blank.
```{r, eval=FALSE}
Fielding %>% 
  filter(playerID == "lemahdj01") %>% 
  arrange(desc(yearID))
```


```{r, echo=FALSE}
Fielding %>% 
  filter(playerID == "lemahdj01") %>% 
  arrange(desc(yearID)) %>% 
  kable() %>% 
  kable_styling("striped") %>% 
  row_spec(1:3, background = "lightblue") %>% 
  row_spec(4:6, background = "skyblue") %>% 
  row_spec(7:9, background = "cornflowerblue") %>% 
  row_spec(14:17, background = "lightblue") %>% 
  row_spec(18:21, background = "skyblue") %>% 
  row_spec(22:25, background = "cornflowerblue") %>% 
  row_spec(26:28, background = "mediumpurple") %>% 
  scroll_box(height = "300px")
```


#### Salaries
The `Salaries` table is very simple with only five variables. One limitation of it is restricted time frame; the earliest season available is 1985 and the latest season is seven years prior to the current year.  
```{r, eval=FALSE}
Salaries %>% 
  filter(yearID == 2016) %>% 
  sample_n(15) 
```

```{r, fig.width=6, echo=FALSE}
Salaries %>% 
  filter(yearID == 2016) %>% 
  sample_n(15) %>% 
  kable(align = "llllr") %>% 
  kable_styling("striped") %>% 
  scroll_box(height = "300px")
```


#### Teams
This is what the `Teams` table looks like for the 2021 season. It includes 48 variables, covering team identifications, standings, totals for numerous statistics, home ballpark, attendance, park factors, and their ID for a couple other databases.
```{r, eval=FALSE}
Teams %>% 
  arrange(desc(yearID)) %>% 
  head(30)
```

```{r, echo=FALSE}
Teams %>% 
  arrange(desc(yearID)) %>% 
  head(30) %>% 
  kable() %>% 
  kable_styling("striped") %>% 
  column_spec(c(4, 41), bold = T) %>% 
  scroll_box(width = "100%", height = "300px")
```


### Visualizations {.tabset}

Here are a few visualizations to help show the data available in the **Lahman** package.

Below is a visualization for each table mentioned above

#### Batting

```{r, fig.align='center', fig.width=7, echo=FALSE}
Batting %>% 
  group_by(yearID) %>% 
  summarize(totalHR = sum(HR)) %>% 
  filter(yearID >= 1990) %>% 
  ggplot(aes(x = yearID, y = totalHR, fill = factor(yearID))) +
  geom_col() +
  labs(title = "Total Home Runs from 1990-2021", x = "Year", y = "Total HRs") +
  theme_bw() +
  theme(legend.position = "none")
```

#### Pitching

```{r, fig.align='center', fig.width=7, echo=FALSE, message=FALSE, warning=FALSE}
Pitching %>% 
  group_by(yearID) %>% 
  filter(yearID >= 1921, !is.na(ERA)) %>% 
  summarize(avgERA = mean(ERA)) %>% 
  ggplot(aes(x = yearID, y = avgERA)) + 
  geom_line(size = 1, color = "dodgerblue") +
  labs(title = "Average ERA in the Last 100 Years",
       x = "Year", y = "Average ERA") +
  theme_bw()
```

#### Fielding

```{r, fig.align='center', fig.width=7, echo=FALSE}
Fielding %>% 
  filter(yearID >= 2012) %>% 
  ggplot(aes(x = factor(POS), y = DP)) +
  geom_jitter(alpha = 0.4, aes(color = factor(yearID)), 
              position = position_jitter(0.2)) +
  labs(title = "Number of Double Plays by Position", 
       x = "Position", y = "Double Plays", color = "Year") +
  theme_bw()
```

#### Salaries

```{r, fig.align='center', fig.width=7, echo=FALSE}
Salaries %>% 
  group_by(yearID) %>% 
  summarize(Low = min(salary), Median = median(salary), 
                     Mean = mean(salary), High = max(salary)) %>% 
  ggplot(aes(x = yearID)) +
  geom_ribbon(aes(ymin = Low, ymax = High), 
                        fill = "lightblue", color = "black", alpha = 0.5) +  
  geom_line(aes(y = Median), color = "black", linetype = "dashed") + 
  geom_line(aes(y = Mean), color = "black", linetype = "dotted") + 
  scale_y_continuous(labels = scales::label_dollar()) + 
  theme_bw() + 
  labs(title = "MLB Salary Range Over Time", 
          subtitle = "Median = Dashed Line, Mean = Dotted Line",       
          y = "Salary", 
          x = "Season")
```

### {-}

```{r, include=FALSE, fig.align='center', fig.width=10, fig.height=10}
a <- Batting %>% 
  group_by(yearID) %>% 
  summarize(totalHR = sum(HR)) %>% 
  filter(yearID >= 1990) %>% 
  ggplot(aes(x = yearID, y = totalHR, fill = factor(yearID))) +
  geom_col() +
  labs(title = "Total Home Runs from 1990-2021", x = "Year", y = "Total HRs") +
  theme_bw() +
  theme(legend.position = "none")
b <- Pitching %>% 
  group_by(yearID) %>% 
  filter(yearID >= 1921, !is.na(ERA)) %>% 
  summarize(avgERA = mean(ERA)) %>% 
  ggplot(aes(x = yearID, y = avgERA)) + 
  geom_line(size = 1, color = "dodgerblue") +
  labs(title = "Average ERA in the Last 100 Years",
       x = "Year", y = "Average ERA") +
  theme_bw()
c <- Fielding %>% 
  filter(yearID >= 2012) %>% 
  ggplot(aes(x = factor(POS), y = DP)) +
  geom_jitter(alpha = 0.4, aes(color = factor(yearID)), 
              position = position_jitter(0.2)) +
  labs(title = "Number of Double Plays by Position", 
       x = "Position", y = "Double Plays", color = "Year") +
  theme_bw()
d <- Salaries %>% 
  group_by(yearID) %>% 
  summarize(Low = min(salary), Median = median(salary), 
                     Mean = mean(salary), High = max(salary)) %>% 
  ggplot(aes(x = yearID)) +
  geom_ribbon(aes(ymin = Low, ymax = High), 
                        fill = "lightblue", color = "black", alpha = 0.5) +  
  geom_line(aes(y = Median), color = "black", linetype = "dashed") + 
  geom_line(aes(y = Mean), color = "black", linetype = "dotted") + 
  scale_y_continuous(labels = scales::label_dollar()) + 
  theme_bw() + 
  labs(title = "MLB Salary Range Over Time", 
          subtitle = "Median = Dashed Line, Mean = Dotted Line",       
          y = "Salary", 
          x = "Season")
(a | b)/(c | d)
```


**Looking Ahead:** Many of these tables contain minimal variables that are specific to one part of baseball (such as pitching or fielding). It would be helpful if we could _join_ multiple tables together.
We will talk more about creating plots in the "Visualizations" section.


