---
title: "Lahman"
output: html_document
date: ''
---

<link rel="preconnect" href="https://fonts.googleapis.com">
<link rel="preconnect" href="https://fonts.gstatic.com" crossorigin>
<link href="https://fonts.googleapis.com/css2?family=Open+Sans:wght@300&family=Source+Sans+Pro&display=swap" rel="stylesheet">
<style>

body{
font-family: 'Source Sans Pro', sans-serif;
}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
```

The **Lahman** package contains several tables with data on pitching, hitting, fielding, and more. This package is updated frequently, so it is possible some tables may change names in a later version. The following references the 10.0-1 version. 

The main tables are:
- `People` : player names, dates of birth, death, and other biographical info
- `Batting` : batting statistics 
  - `battingStats` : calculates batting average (BA) plate appearances (PA), total bases (TB), slugging percentage (SlugPct), on-base percentage (OBP), on-base percentage + slugging (OPS), and batting average on balls in play (BABIP)
- `Pitching` : pitching statistics 
- `Fielding` : fielding statistics 

A full list of the tables and their contents can be found in the [documentation](https://cran.r-project.org/web/packages/Lahman/Lahman.pdf).

**type of data available:** summary statistics, broken down season-by-season(by team for `Batting` / `Pitching`, by team _and_ position for `Fielding`)
FIND A WAY TO REPHRASE - BETTER PLACEMENT

```{r, message=FALSE, warning=FALSE}
library(Lahman)
```


should have a way to show the variables in each table -> skim would probably be too long, maybe use head?

```{r}
kable(head(People)) %>% 
  kable_styling("striped") %>% 
  scroll_box(width = "100%")
```

should we do this for each of the four major tables?

```{r, fig.align='center'}
Batting %>% 
  group_by(yearID) %>% 
  summarize(totalHR = sum(HR)) %>% 
  filter(yearID >= 1996) %>% 
  ggplot(aes(x = yearID, y = totalHR, fill = factor(yearID))) +
  geom_col() +
  labs(x = "Year", y = "Total HRs") +
  theme_bw() +
  theme(legend.position = "none")
```


```{r, fig.align='center'}
Pitching %>% 
  group_by(yearID) %>% 
  filter(yearID >= 1921, !is.na(ERA)) %>% 
  summarize(avgERA = mean(ERA)) %>% 
  ggplot(aes(x = yearID, y = avgERA)) + 
  geom_line(size = 1, color = "dodgerblue") +
  labs(title = "Average ERA in the last 100 years",
       x = "Year", y = "Average ERA") +
  theme_bw()
```


what graph for Fielding?

```{r, fig.align='center'}
Fielding %>% 
  filter(yearID >= 2012) %>% 
  ggplot(aes(x = factor(POS), y = DP)) +
  geom_jitter(alpha = 0.4, aes(color = factor(yearID)), 
              position = position_jitter(0.2)) +
  labs(title = "The Number of Double Plays by Position", 
       x = "Position", y = "Double Plays", color = "Year") +
  theme_bw()
```


_Notes_

- include the salary table?
- what visualizations should be included?
- am I allowed to say "you" or "we"?
