---
title: "Lahman"
output: html_document
date: ''
---

<link rel="preconnect" href="https://fonts.googleapis.com">
<link rel="preconnect" href="https://fonts.gstatic.com" crossorigin>
<link href="https://fonts.googleapis.com/css2?family=Open+Sans:wght@300&family=Source+Sans+Pro&display=swap" rel="stylesheet">
<style>

body{
font-family: 'Source Sans Pro', sans-serif;
}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
```

The **Lahman** package contains several tables with data on pitching, hitting, fielding, and more. This package is updated with each season. The following references the 10.0-1 version. 

**type of data available:** season-by-season, broken down by team for `Batting` / `Pitching` (so if a player played on multiple teams in one season, each team as their own row), broken down by team _and_ position for `Fielding`

broken down by team for players that changed teams during a season and position for fielding (find a way to illustrate this tables) WORK ON!!!!!

The main tables are:

- `People` : player names, dates of birth, death, and other biographical info
- `Batting` : batting statistics 
  - `battingStats` : calculates batting average (BA) plate appearances (PA), total bases (TB), slugging percentage (SlugPct), on-base percentage (OBP), on-base percentage + slugging (OPS), and batting average on balls in play (BABIP)
- `Pitching` : pitching statistics 
- `Fielding` : fielding statistics 
- `Salaries` : player name, year, team, league, and salary (from 1985-2016)
- `Teams` : overall season statistics
- Additional tables cover information such as All Star appearances, post season data, managers, Hall of Fame voting, awards, parks, and colleges.

A full list of the tables and their contents can be found in the [documentation](https://cran.r-project.org/web/packages/Lahman/Lahman.pdf).


```{r, message=FALSE, warning=FALSE}
library(Lahman)
```


#### People
```{r}
People %>% 
  sample_n(6) %>% 
  kable() %>% 
  kable_styling("striped") %>% 
  #column_spec(column = 6, width = "125px") %>% 
  scroll_box(width = "100%")
```

?: is there a way to make the text not wrap (without having to individually adjust every column width)? 


#### Batting 
Example: **José Iglesias**
```{r}
# People %>% filter(nameFirst == "Mookie", nameLast == "Betts")
# People %>% filter(nameFirst == "Jose", nameLast == "Iglesias")

iglesias <- Batting %>% 
  filter(playerID == "iglesjo01") %>% 
  arrange(desc(yearID))

kable(iglesias) %>% 
  kable_styling("striped") %>% 
  row_spec(1:2, background = "skyblue") %>% 
  row_spec(9:10, background = "lightseagreen") %>% 
  scroll_box(height = "300px")
```


`battingStats()` produces a data frame with the same variables as `Batting`, and additionally calculates some statistics (as mentioned above). 

```{r}
iglesias_stats <- battingStats(iglesias) %>% 
  dplyr::select(playerID, yearID, teamID, BA:BABIP)

kable(iglesias_stats) %>% 
  kable_styling("striped") %>% 
  row_spec(1:2, background = "skyblue") %>% 
  row_spec(9:10, background = "lightseagreen") %>% 
  scroll_box(height = "300px")
```


#### Pitching 
Example: **Justin Verlander**
```{r}
verlander <- Pitching %>% 
  filter(playerID == "verlaju01") %>% 
  arrange(desc(yearID))

kable(head(verlander)) %>% 
  kable_styling("striped") %>% 
  row_spec(4:5, background = "lightblue") %>% 
  scroll_box(width = "100%")

# Batting %>% 
#   filter(playerID == "scherma01")
```


#### Fielding
Example: **DJ LeMahieu**
```{r}
Fielding %>% 
  filter(playerID == "lemahdj01") %>% 
  arrange(desc(yearID)) %>% 
  kable() %>% 
  kable_styling("striped") %>% 
  scroll_box(height = "300px")
```


#### Salaries
```{r}
Salaries %>% 
  filter(yearID == 2016) %>% 
  sample_n(15) %>% 
  kable(align = "llllr") %>% 
  kable_styling("striped") %>% 
  scroll_box(height = "300px")
```


#### Teams
```{r}
Teams %>% 
  arrange(desc(yearID)) %>% 
  head(30) %>% 
  kable() %>% 
  kable_styling("striped") %>% 
  column_spec(c(4, 41), bold = T) %>% 
  scroll_box(width = "100%", height = "300px")
```


```{r, fig.align='center'}
Batting %>% 
  group_by(yearID) %>% 
  summarize(totalHR = sum(HR)) %>% 
  filter(yearID >= 1996) %>% 
  ggplot(aes(x = yearID, y = totalHR, fill = factor(yearID))) +
  geom_col() +
  labs(x = "Year", y = "Total HRs") +
  theme_bw() +
  theme(legend.position = "none")
```

can adjust figure width
```{r, fig.align='center'}
Pitching %>% 
  group_by(yearID) %>% 
  filter(yearID >= 1921, !is.na(ERA)) %>% 
  summarize(avgERA = mean(ERA)) %>% 
  ggplot(aes(x = yearID, y = avgERA)) + 
  geom_line(size = 1, color = "dodgerblue") +
  labs(title = "Average ERA in the last 100 years",
       x = "Year", y = "Average ERA") +
  theme_bw()
```

show in table players who played multiple positions (utility player)
what graph for Fielding?

```{r, fig.align='center'}
Fielding %>% 
  filter(yearID >= 2012) %>% 
  ggplot(aes(x = factor(POS), y = DP)) +
  geom_jitter(alpha = 0.4, aes(color = factor(yearID)), 
              position = position_jitter(0.2)) +
  labs(title = "The Number of Double Plays by Position", 
       x = "Position", y = "Double Plays", color = "Year") +
  theme_bw()
```

This is definitely not correct
```{r, fig.align='center'}
Salaries %>% 
  filter (yearID >= 2012) %>% 
  ggplot(aes(x = yearID)) +
  #geom_ribbon(aes(ymin = 0, ymax = salary)) +
  geom_ribbon(aes(ymin = min(salary), ymax = max(salary))) +
  #geom_ribbon(aes(ymin = salary - 1, ymax = salary + 1)) +
  theme_bw()
```


graph showing players

salary - geom_ribbon
scales package to convert it to dollars


_Notes_

- help with geom_ribbon and geom_label (figured out geom_label in fangraphs file)
- would battingLabels data frame be useful?
