---
title: "Modeling"
output: html_document
date: ''
---

<link rel="preconnect" href="https://fonts.googleapis.com">
<link rel="preconnect" href="https://fonts.gstatic.com" crossorigin>
<link href="https://fonts.googleapis.com/css2?family=Open+Sans:wght@300&family=Source+Sans+Pro&display=swap" rel="stylesheet">
<style>

body{
font-family: 'Source Sans Pro', sans-serif;
}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

What do we want to be modeling?

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(Lahman)
```


### Linear Model

We will use data from the Lahman package.
```{r}
bat <- battingStats() %>% 
  filter(yearID >= 2012)
```

The `lm()` function is used to fit linear models. It requires a formula, following the format: response ~ predictor(s).

#### One Variable

```{r}
model1 <- lm(R ~ PA, data = bat)
summary(model1)
```

```{r, fig.align='center', fig.width=7}
bat %>% 
  ggplot(aes(x = PA, y = R)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm") +
  theme_bw()
```

#### Multiple Variables

Our goal is to model runs using plate appearances, batting average, and on-base plus slugging. The first step is removing any NA values from those variables.

```{r}
bat <- bat %>% 
  filter(!is.na(PA), !is.na(BA), !is.na(OPS))
```

Now we can build the model. 

```{r}
model2 <- lm(R ~ PA + BA + OPS, data = bat)
model2
summary(model2)
```


### Predicting Winning Percentage
(pg 95 of _Analyzing Baseball Data with R_)

For this section we will be using the `Teams` table. 

```{r}
team_pred <- Teams %>% 
  filter(yearID >= 2012)
```


Say we want to predict winning percentage. To do this we need to create two new variables: run differential (RD) and winning percentage (WPCT). Run differential is calculated with runs scored (R) and opponents runs scores (RA). Winning percentage needs the amount of wins and losses. 

```{r}
team_pred <- team_pred %>% 
  mutate(RD = R - RA,
         WPCT = W / (W + L))
```

Now it's time to build the model.
```{r}
model3 <- lm(WPCT ~ RD, data = team_pred)
model3
summary(model3)
```

```{r, fig.align='center', fig.width=7}
team_pred %>% 
  ggplot(aes(x = RD, y = WPCT)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm") +
  theme_bw()
```







